The problem and promise of scale in multilayer networks

### 2020-01-23 paper/manuscript.Rmd
add perspective

>>>>

Social network positions are frequently used to represent the social environment that an individual experiences or their social phenotype, and they are increasingly linked to important fitness outcomes for individuals among species ranging from primates (e.g. Brent, Ruiz-Lambides, & Platt, 2017; Thompson, 2019) to dolphins (e.g. Stanton & Mann, 2012; Stanton, Gibson, & Mann, 2011) to birds (e.g. McDonald, 2007; Royle, Pike, Heeb, Richner, & Kolliker, 2012). But there are many ways that scale comes into play in determining how characteristics of an individual’s social network position are related to fitness that depend on the hypotheses being tested and logistical constraints. The two scales that repeatedly come across as important for understanding fitness consequences are temporal and social scale (Almeling, Hammerschmidt, Sennhenn-Reulen, Freund, & Fischer, 2016; Berger, Lemaître, Allainé, Gaillard, & Cohas, 2015; Brent et al., 2017).

Social network structures depend on the time frame used to create them, which can make different networks difficult to compare (Castles et al., 2014). Furthermore, social network structures can vary by season and indicate social environments that an individual experiences (e.g. Brent, MacLarnon, Platt, & Semple, 2013; Hamede, Bashford, McCallum, & Jones, 2009; Holekamp, Smith, Strelioff, Van Horn, & Watts, 2012), or they can be consistent across time, which can be used to reflect social phenotypes (Aplin et al., 2015; Stanley, Mettke-Hofmann, Hager, & Shultz, 2018). Seasonal fluctuations have the potential to have disease transmission and fitness effects for individuals (Balasubramaniam et al., 2019; Chen et al., 2014).

Additionally, in long-lived species especially, social network metrics of individuals can change over their lifetimes in ways that can have impacts on their reproductive success and longevity (Berger et al., 2015; Brent et al., 2017). These findings highlight the necessity to think carefully about the temporal scale that is meaningful for the proposed hypotheses, especially for understanding individual social phenotypes or environments and how they relate to that individual’s fitness. Further complicating the issue, social network metrics can change over stages of ontogeny, but individuals progress through development at different rates (Tarka, Guenther, Niemelä, Nakagawa, & Noble, 2018). This variation can make defining appropriate temporal scales for testing the fitness implications of an individual’s social network position over different stages of ontogeny even more difficult. However, thinking about temporal scales at such a fine level can provide insights for important stages of development that may have otherwise been missed (CITE-myself??). However, multilayer networks provide a framework to test hypotheses about how social dynamics over time relate to fitness. Because of the non-independent nature of social interactions, especially over time, it is hard to test hypotheses about social dynamics with a suitable null control (Farine, 2017; Proskurnikov & Tempo, 2017). Using the framework of multilayer networks, making layers of different temporal periods provides a way to make appropriate null models to test questions about the dynamics of an individual’s social network position over multiple aggregated time periods.

Social scale is another major consideration for social network analysis, especially when thinking about different types of interactions and associations in the same analysis (Carter, Lee, & Marshall, 2015; Castles et al., 2014; Farine, 2015). It is a common trope that individuals need to be associating in order to interact and an assumption that proximity is proxy for interacting (Farine, 2015). For instance, baboons need to be in proximity in order to groom each other, but this trope is only thinking about interactions that humans are most prone to seeing and valuing with our bias toward visual perception. Animals have certain interactions that do not require proximity such as long-distance vocalization and olfactory signals (Carter et al., 2015). With improved and more affordable technology, we can now start to understand some of these interactions that do not require proximity, particularly long-distance communication networks with devices like microphone arrays (Snijders & Naguib, 2017). The multilayer network framework further enables researchers to look at these different social scales and test how good a proxy proximity is for various forms of interacting in a holistic way.

These different social scales can have different fitness repercussions for individuals. Rhesus macaques (*Macaca mulatta*) that had stronger, more stable associates and grooming partners but not more partners in general had greater survivorship than those who did not (Ellis, Snyder-Mackler, Ruiz-Lambides, Platt, & Brent, 2019). This finding highlights the need to look at social network positions at both multiple temporal and social scales to understand the social positions and fitness repercussions of individuals, and multilayer networks provide an avenue to do so considering the whole social system rather than just one type of social interaction at a time, which is what has been done under most circumstances but is not very realistic.

Multilayer networks further provide unique challenges and opportunities for understanding the social phenotypes and environments of individuals and their fitness outcomes. Picking the proper temporal scale is important but a challenge because all layers in the multilayer network should be made for the same time period, and the period needs to be long enough for each layer to have enough data for robust networks (Farine, 2017). These considerations may be limiting factors when answering questions that require timeframes that are relevant to different individuals. On the other hand, multilayer network metrics provide an opportunity to help simplify statistical models. Interactions that happen mere meters apart, proximity associations, and long-distance interactions can be encapsulated in one network to gain a single metric to describe the various social scales of an individual’s network position. This multilevel network metric would be a meaningful simplification of many network metrics that could help simplify statistical analyses and deal with the frequent correlations between social network metrics in different types of networks (Castles et al., 2014). This method could be an alternative to using PCA to simplify many social network metrics into one social phenotype (e.g. Foster et al., 2012; Oldham et al., 2019). Future analyses should explore the differences in how well each of these methods describe and predict fitness outcomes of the complex social network positions of individuals.
